---
title: "Extract vertical profile data and export to a single CSV file"
author: "Peter Desmet"
output:
  html_document:
    toc: yes
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, results="hide", warning=FALSE, message=FALSE}
library(bioRad)
library(dplyr)
source("vp_to_df.R")
```

## 1. Define extraction settings

Since the data we want to include for the TIMAMP visualization might be different from what we have downloaded (or already stored locally), we need to define some extraction settings/filters. Those include date range, countries and radars to include and where the vertical profile data are stored locally (a file path relative to this script):

```{r extraction_settings}
extraction_settings <- list()
extraction_settings[["start_date"]]   <- "2016-10-03"
extraction_settings[["end_date"]]     <- "2016-10-03"
extraction_settings[["countries"]]    <- c(
  "se"
)
extraction_settings[["radars"]]       <- c(
  "ang", "arl", "ase", "hud", "kir", "kkr", "lek", "lul", "osu", "ovi", "var", "vil"
)
extraction_settings[["raw_data_dir"]] <- "../data/raw/se-sample/" # Needs a trailing slash!
```

## 2. Create a list of data file paths

Build a list of all the file (paths) that meet our extraction criteria:

```{r file_paths_1}
vp.file_paths <- bioRad::retrieve_vp_paths(
  start_date = extraction_settings$start_date,
  end_date = extraction_settings$end_date,
  country = extraction_settings$countries,
  radar = extraction_settings$radars,
  path = extraction_settings$raw_data_dir
)
```

The file paths only start at the level of the data directory, so we need to append those with the path to the data directory itself:

```{r file_paths_2}
vp.file_paths <- paste0(extraction_settings$raw_data_dir, vp.file_paths)
head(vp.file_paths)
```

## 3. Read data

### Single file example

Read a single vertical profile file (vp) with [bioRad](https://github.com/adokter/bioRad):

```{r single_read}
vp.single <- bioRad::readvp("../data/raw/se-sample/2016/10/03/20/00/searl_vp_20161003T2000Z.h5")
```

Have a look at the structure:

```{r single_summary}
summary(vp.single)
```

And the data:

```{r single_data}
vp.single$data
```

Create a data.frame with the data and metadata we need (using the `vp_to_df` function):

```{r single_df}
head(vp_to_df(vp.single))
```

### All files

Now, let's do this for all files we selected with our extraction settings. Get file paths:

```{r all_read}
vp.some <- bioRad::readvp.list(c(
  "../data/raw/se-sample/2016/10/03/20/00/searl_vp_20161003T2000Z.h5",
  "../data/raw/se-sample/2016/10/03/20/15/searl_vp_20161003T2015Z.h5",
  "../data/raw/se-sample/2016/10/03/20/30/searl_vp_20161003T2030Z.h5",
  "../data/raw/se-sample/2016/10/03/20/45/searl_vp_20161003T2045Z.h5"))
vp.all <- bioRad::readvp.list(vp.file_paths)
```

_Note: bioRad will warn that not all vertical profiles are from a single radar, but that's fine._

Create a data.frame with the data and metadata we need from all those files:

```{r all_to_df}
vpdf.list = list()
for (i in seq_along(vp.all)) {
  vpdf.list[[i]] <- vp_to_df(vp.all[[i]])
}

vpdf = dplyr::bind_rows(vpdf.list)
```

## 4. Sort data

```{r all_sort}
vpdf <- dplyr::arrange(vpdf, radar_id, date_time, HGHT)
```

```{r all_head}
head(vpdf)
```

## 5. Export data as CSV file

Note that `NaN` values will be treated like `NA` values.

```{r all_export}
write.csv(vpdf, file = "../data/interim/vp.csv", na = "", row.names = FALSE)
```
