# Download vp data {#download_vp}

```{r 2_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This chapter describes how to download vp data from the [ENRAM bird profile data repository](http://enram.github.io/data-repository). If you already have all the vp data (= h5 files) you need on your computer, you can skip this completely. ðŸ‘‹

## Define download settings

To download specific vp data from the [ENRAM bird profile data repository](http://enram.github.io/data-repository), we need to specify a number of settings/filters, such as:

* date range
* countries to include
* radars to include (unless you only want to select on whole countries)
* where to download the vp data to (= a file path relative to this file)

All these are stored in `download_settings`:

```{r 2_download_settings}
download_settings <- list()
download_settings[["start_date"]]   <- "2017-03-01"
download_settings[["end_date"]]     <- "2017-04-30"
download_settings[["countries"]]    <- c("se")
download_settings[["radars"]]       <- c("ang", "kkr")
download_settings[["raw_data_dir"]] <- "../data/raw/"
```

## Download vp data

```{r 2_libraries, results="hide"}
library(bioRad)
```

Next, we download the data. You better get some â˜• as this might take a while.

```{r 1_download_data, results="hide", eval=FALSE}
bioRad::download_vp(
  start_date = download_settings$start_date,
  end_date = download_settings$end_date,
  country = download_settings$countries,
  radar = download_settings$radars,
  localpath = download_settings$raw_data_dir
)
```

Finished? Great! ðŸ‘ In the next chapter we'll see how to export the relevant data from all those h5 files to a single CSV file.
