# Extract vp data {#extract-vp}

```{r extract_rmd_settings, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

In this chapter we'll select, read, combine, filter and export data from vp files to a single, unaggregated CSV file.

```{r extract_load_libraries, results = "hide"}
library(bioRad)
library(dplyr)
library(loopr)
library(lubridate)
source("functions/vp_to_df.R")
source("functions/load_settings.R")
```

## Load settings

All of the filtering options we'll use here are defined in **yaml settings file**, which we pointed to in `setup.Rmd` (see \@ref(settings)). Let's load those:

```{r extract_load_settings}
settings <- load_settings(settings_file, radars_metadata_file)
```

Found **`r length(settings$general$radar_ids_5char)` radars from `r length(settings$general$countries)` countries** in the settings.

## Select date range & radars

Select vp files on the date range and radars we defined in our settings (_file selection_):

```{r extract_get_file_paths}
vp_file_paths <- bioRad::retrieve_vp_paths(
  start_date = settings$general$start_date,
  end_date = settings$general$end_date,
  radar = settings$general$radar_ids_3char,
  path = raw_data_dir
)
```

The file paths we get back only start from the data directory, so we need to append those with the path to the data directory itself:

```{r extract_complete_file_paths}
vp_file_paths <- paste0(raw_data_dir, vp_file_paths)
```

There are **`r length(vp_file_paths)` vp files** that meet our criteria. Preview:

```{r extract_preview_file_paths}
head(vp_file_paths, 5)
```

## Read vp files

We read all those files with bioRad. This could take a while (☕ or 🛌 time):

```{r extract_read_vp_data_take_a_break}
vp_files <- bioRad::readvp.list(vp_file_paths)
```

## Select variables

A single vp file contains a data frame with heights as rows and variables as columns:

```{r extract_show_vp_data_structure}
str(vp_files[[1]]$data)
```

We only need a couple of variables (`HGHT`, `u`, `v`, `dens`, `dd`, `ff`) and attributes (`date_time` and `radar_id`) from the vp_files (_column selection_). We retrieve those with the custom function `vp_to_df()` after which we combine all data in one single data frame, sorted on `radar_id`, `date_time`, and `HGHT`:

```{r extract_select_columns_and_sort}
vp_data <- list()
variables = c("u", "v", "dens", "dd", "ff") # HGHT is added by default in vp_to_df()
for (i in seq_along(vp_files)) {
  vp_data[[i]] <- vp_to_df(vp_files[[i]], variables)
}

dplyr::bind_rows(vp_data) %>%
dplyr::arrange(radar_id, datetime, HGHT) -> vp_data
```

To uniquely reference records later, we add an `id` column:

```{r extract_add_id_column}
vp_data <- cbind("id" = seq.int(nrow(vp_data)), vp_data)
```

That data frame contains **`r nrow(vp_data)` records**.

## Calculate MTR per height

We also want migration traffic rate (mtr) per height for further analysis, so we calculate and add it as an extra column:

```{r extract_calculate_mtr}
vp_data$mtr <- vp_data$ff * vp_data$dens / 5

# Add our new mtr to variables
variables <- c(variables, "mtr")
```

Preview

```{r extract_vp_data_preview}
head(vp_data, 20)
```

## Add day/night information

To add day/night information, we first create a sunrise/sunset dataframe for each radar/date combination, using the latitude/longitude from our radar metadata and the `suntime()` function in bioRad:

```{r extract_get_day_and_night}
# Create a simple dataframe from the settings with radar, latitude, longitude
bind_rows(lapply(settings$radars, data.frame)) %>%
select(radar_id, latitude, longitude) -> radars_lat_long

# Group vp_data by radar and date
vp_data %>%
mutate(date = as.Date(datetime)) %>%
group_by(radar_id, date) %>%
summarize() %>%

# Join with radar_lat_long to get the latitude and longitude column
left_join(radars_lat_long, by = "radar_id") %>%

# Use bioRad to add sunrise/sunset information
mutate(sunrise = bioRad::suntime(date = date, lat = latitude, lon = longitude, rise = TRUE)) %>%
mutate(sunset = bioRad::suntime(date = date, lat = latitude, lon = longitude, rise = FALSE)) -> radars_dates_sunriset

# Preview
head(radars_dates_sunriset, 5)
```

We then combine this sunrise/sunset information back with our vp data, to figure out if it is day or night and to which date they belong (`date_of_sunset`). For the latter, we consider night timestamps between midnight and sunrise as belonging to the previous date:

```{r extract_calculate_sunriset}
vp_data %>%
# A date (not datetime) column to match with radars_sunriset
mutate(date = as.Date(datetime)) %>%

# Combine vp data with radar_sunriset information
right_join(radars_dates_sunriset, by = c("radar_id" = "radar_id", "date" = "date")) %>%
  
# Define nights as starting before sunrise and after sunset
mutate(day_night = if_else(
  datetime < sunrise | datetime > sunset,
  "night",
  "day"
)) %>%
  
# Calculate the date of sunrise:
# For days: keep the date of the datetime
# For nights: rewind the time with 12 hours, so that night timestamps between
#             midnight and sunrise are considered belonging to previous day.
mutate(date_of_sunset = if_else(
  day_night == "night",
  datetime - hours(12),
  datetime
)) %>%
mutate(date_of_sunset = format(date_of_sunset, format = "%Y%m%d")) %>%

# Remove unneeded columns
select(-date, -latitude, -longitude, -sunrise, -sunset) -> vp_data
```

Preview:

```{r extract_calculate_sunriset_preview}
head(vp_data, 20)
```

## Filter out heights

For each radar we only want to select the data above 200m above ground level (AGL) (_row selection_). Since the heights in the data are expressed in above sea level (ASL), the height range to exclude is the _altitude of the radar + 100m_, which differs from radar to radar. Those height ranges are defined in our settings (see \@ref(settings)).

To do this, we take the `include_heights` range for each radar[^2] and set the data outside that range to `NA`.

[^2]: For a radar the `include_heights` range can contain a `min_height`, `max_height`, both or none for a radar. If the value is something other than an integer, the general `min_height` or `max_height` is used instead.

```{r extract_filter_heights}
for (radar in settings$radars) { # For each radar
  # Create height subset for that radar
  subset = 
    vp_data %>%
    filter(
      radar_id == radar$radar_id & # Specific radar
      (HGHT < radar$min_height |   # Below min heignt
       HGHT > radar$max_height)    # Above max height
    ) %>%
    # Set variables to NA
    mutate(u = NA, v = NA, dens = NA, dd = NA, ff = NA, mtr = NA) %>%
    # Mention that the row was excluded because of height
    mutate(exclusion_reason = "height")
  
  # Insert subset back into vp_data
  insert(vp_data, subset, by = "id") -> vp_data
}
```

Preview:

```{r extract_filter_heights_preview}
head(vp_data, 20)
```

## Filter out datetimes

For some radars certain datetime ranges need to be excluded due to (remaining) rain clutter (_row selection_). Those datetime ranges were discovered during manual inspection and are defined in our settings (see \@ref(settings))._

To do this, we take the `exclude_datetimes` range for each radar and set the data inside that range to `NA`.

```{r extract_filter_datetimes}
vp_data_temp <- vp_data
for (radar in settings$radars) { # For each radar
  for (i in seq_along(radar$exclude_datetime_from)) { # For each exclude_datetime
    # Create datetime subset for that radar
    subset = 
      vp_data_temp %>%
      filter(
        radar_id == radar$radar_id & # Specific radar
        between(
          datetime,
          radar$exclude_datetime_from[[i]], # From this datetime
          radar$exclude_datetime_to[[i]]    # Until this datetime
        )
      ) %>%
      # Set variables to NA
      mutate(u = NA, v = NA, dens = NA, dd = NA, ff = NA, mtr = NA) %>%
      # Mention that the row was excluded because of datetime
      # This will overwrite any previous "height" reason
      mutate(exclusion_reason = "datetime")
      
    # Insert subset back into vp_data
    insert(vp_data, subset, by = "id") -> vp_data
  }
}
```

Preview:

```{r extract_filter_datetimes_preview}
head(vp_data, 20)
```

## Export to a CSV file

Finally, we export the data to a CSV file[^3]:

```{r extract_vp_export}
# Remove id column
vp_data %>% select(-id) -> vp_data

# Write data to file (filename is dynamically created by load_settings)
write.csv(vp_data, file = paste0(processed_data_dir, settings$general$vp_output_file), na = "", row.names = FALSE)
```

[^3]: The data frame `vp_data` contains `NaN` and `NA` values, [which have a different meaning](https://github.com/enram/vp-processing/issues/10#issuecomment-302235736). That difference gets lost in the CSV file: all are treated as blank (= `NA`) values, which is fine for visualizations. If you want to keep that difference, you need work further with `vp_data`.
