# Getting started

```{r 1_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

_Note: This documentation assumes you are working in RStudio._

## Install

In this tutorial we use a number of R packages, which you need to install first:

```{r 1_install_packages, eval=FALSE}
install.packages(c(
  "devtools",  # To install packages from GitHub
  "dplyr",     # To process dataframes
  "lubridate", # To process dates
  "bookdown"   # To generate this documentation
))
devtools::install_github("adokter/bioRad", ref="master") # To process vp data
```

## Download the code

To use and especially _adapt_ the code of this tutorial, fork or clone this repository: https://github.com/enram/timamp-etl

## Edit the code

Once downloaded, go to the `/src` directory:

1. Open `timamp-etl.Rproj` as a project in RStudio
2. Open one of the `.Rmd` files in RStudio to see and edit the code.
3. Test your code by running it.

## Generating fancy documentation 📕

What your are reading here (in your browser) is all generated from the `.Rmd` files in the `src` directory, using the R package [bookdown](https://bookdown.org/yihui/bookdown/).

In fact, **if you notice a typo in this documentation**, you can correct it by clicking the pencil icon in the menu at the top, which will take you straight to the correct `.Rmd` file in the [original repository on GitHub](https://github.com/enram/timamp-etl). If you click the pencil icon there, it will be able to edit right away or suggest to create a fork (= copy) first, depending on your access rights.

If you want to generate fancy documentation for your own adapted code, then:

1. Go to `Build` pane in RStudio
2. Click the `Build book` button
3. Wait for it the build to finish: that might take a while, as it will run all the `.Rmd` files.
4. Go to the `docs` repository
5. Open `index.html` in your browser
6. See your updated documentation

For more info on debugging or hosting your documentation (i.e. on GitHub pages), see [this bookdown book on bookdown](https://bookdown.org/yihui/bookdown/). 😊
