# Process vp data for the bird migration flow visualization {#vp-to-flowviz}

```{r flowviz_rmd_settings, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

In this chapter we'll process the vp data (CSV file) and metadata for the [bird migration flow visualization](http://enram.github.io/bird-migration-flow-visualization/viz/).

```{r flowviz_load_libraries, results = "hide"}
library(bioRad)
library(dplyr)
library(lubridate)
library(circular)
source("functions/load_settings.R")
```

## Load settings

Load the same settings we defined and used before (see \@ref(settings)):

```{r flowviz_load_settings}
settings <- load_settings(settings_file, radars_metadata_file)
```

Found **`r length(settings$general$radar_ids_5char)` radars from `r length(settings$general$countries)` countries** in the settings.

## Read data from CSV file

To allow this chapter to run without having to process all the vp data, we'll load the data from the CSV file(s) rather than from the dataframe created in \@ref(extract-vp).

Search in the defined processed data directory for csv files:

```{r flowviz_get_csv_files}
processed_data_dir_no_slash <- substr(processed_data_dir, 1, nchar(processed_data_dir)-1) 
csv_files <- dir(processed_data_dir_no_slash, pattern = "\\.csv", full.names = TRUE, recursive = FALSE)
csv_files
```

Read and concatenate files:

```{r flowviz_read_csv_files}
csv_data <- lapply(csv_files, read.csv)
flowviz_data <- bind_rows(csv_data)
```

Preview:

```{r flowviz_preview}
head(flowviz_data)
```

## Add time bins

Add time bins per hour:

```{r flowviz_add_time_bins}
flowviz_data %>%
  mutate(datetime_bin = floor_date(as.POSIXct(datetime, tz = "UTC"), "hour")) -> flowviz_data
```

## Add height bins

Add two height bins: `200-2000` and `above 2000`:

```{r flowviz_add_height_bins}
flowviz_data %>%
mutate(height_bin = case_when(
  .$HGHT >= 200 & .$HGHT < 2000 ~ "1",
  .$HGHT >= 2000 ~ "2"
)) -> flowviz_data
```

Preview:

```{r flowviz_preview_bins}
head(flowviz_data, 20)
```

## Filter out excluded data

Only keep data that should not be excluded:

```{r flowviz_filter_excluded}
flowviz_data %>%
filter(exclusion_reason == "") -> flowviz_data
```

## Aggregate data

Aggregate the data by datetime and height bin, taking the mean for every value. For the direction `dd` we need to calculate the circular mean, with a little helper function:

```{r flowviz_circular_mean}
circular_mean_degrees <- function(directions) {
  # If all directions are NA, mean.circular will throw an error, so we test it beforehand:
  if (all(is.na(directions))) {
    circular_mean = NA
  } else {
    circular_mean <- mean.circular(circular(directions, units = "degrees"), na.rm = TRUE)[[1]]
    # Convert negative to positive degrees (e.g. -45 = 315):
    circular_mean <- if (circular_mean < 0) 360 + circular_mean else circular_mean
  }
  return(circular_mean)
}
```

Aggregate data:

```{r flowviz_aggregate}
flowviz_data %>%
group_by(radar_id, datetime_bin, height_bin) %>%
summarize(
  avg_u = mean(u, na.rm = TRUE),
  avg_v = mean(v, na.rm = TRUE),
  avg_dens = mean(dens, na.rm = TRUE),
  avg_dd = circular_mean_degrees(dd),
  avg_ff = mean(ff, na.rm = TRUE),
  avg_mtr = mean(mtr, na.rm = TRUE)
) %>%
ungroup() -> flowviz_data
```

Preview:

```{r flowviz_preview_aggregate}
head(data.frame(flowviz_data))
```

## Filter out empty rows

Only keep data where at least one of the calculated values is not `NA`:

```{r flowviz_filter_not_na}
flowviz_data %>%
  filter(
    !is.na(avg_u) | 
    !is.na(avg_v) | 
    !is.na(avg_dens) | 
    !is.na(avg_dd) | 
    !is.na(avg_ff) |
    !is.na(avg_mtr)
  ) -> flowviz_data

```

Preview:

```{r flowviz_preview_not_na}
head(data.frame(flowviz_data))
```

## Rename and select columns

Rename and select the columns required by the flowviz:

```{r flowviz_rename}
flowviz_data %>%
rename(
  interval_start_time = datetime_bin,
  altitude_band = height_bin,
  avg_u_speed = avg_u,
  avg_v_speed = avg_v
) %>%
# Use NA for all non-finite values (e.g. NaN -> NA):
mutate(avg_u_speed = ifelse(is.finite(avg_u_speed), avg_u_speed, NA)) %>% 
mutate(avg_v_speed = ifelse(is.finite(avg_v_speed), avg_v_speed, NA)) %>%
select(radar_id, interval_start_time, altitude_band, avg_u_speed, avg_v_speed) -> flowviz_data
```

Preview:

```{r flowviz_preview_rename}
head(flowviz_data)
```

## Export to a CSV file

Export the data to a flowviz data file:

```{r}
write.csv(flowviz_data, file = paste0(processed_data_dir, "flowviz/birds.csv"), na = "", row.names = FALSE)
```

## Create metadata

Not yet implemented, currently created manually (see [example](https://github.com/enram/bird-migration-flow-visualization/blob/master/viz/2/nl-be/radars.json)).
